---
title: "Marketing Mix Modeling: Exploratory Data Analysis"
subtitle: "BrickCraft Studios - Synthetic Marketing Data"
author: "Generated by Databot"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: false
    theme: cosmo
    embed-resources: true
execute:
  warning: false
  message: false
---

::: {.callout-note}
This report was generated using AI under general human direction. At the time of generation, the contents have not been comprehensively reviewed by a human analyst.

<!--
To indicate human review: Delete the line above about contents not being reviewed, and replace this comment with:
The contents have been reviewed and validated by [Your Name], [Your Role] on [Date].
-->
:::

# Introduction

This report presents an exploratory data analysis (EDA) of synthetic marketing and sales data for BrickCraft Studios. The analysis examines the relationships between various marketing channels (TV, Radio, Display Ads, Search Ads, and Social Media) and sales revenue to support Marketing Mix Modeling (MMM).

The dataset contains weekly observations of marketing spend across multiple channels and corresponding sales revenue. This EDA will help identify patterns, relationships, and data quality issues before building predict
 Structure

```{r setup}
library(tidyverse)
library(scales)
library(corrplot)
library(knitr)
library(kableExtra)

# Set theme for plots
theme_set(theme_minimal(base_size = 12))
```

```{r load-data}
# Load the synthetic marketing-sales data
data <- read_csv("data/synthetic-marketing-sales.csv")

# Display first few rows
head(data) |>
  kable(caption = "First 6 rows of the dataset") |>
  kable_styling(bootstrap_options = c("striped", "hover"))
```

## Dataset Overview

```{r data-structure}
# Data dimensions
cat("Dataset dimensions:", nrow(data), "rows ×", ncol(data), "columns\n\n")

# Column structure
glimpse(data)
```

```{r date-range}
# Date range
cat("Date range:", as.character(min(data$date)), "to", as.character(max(data$date)), "\n")
cat("Number of weeks:", n_distinct(data$week), "\n")
cat("Number of years:", n_distinct(data$year), "\n")
```

# Univariate Analysis

## Sales Revenue

```{r sales-summary}
# Sales revenue summary statistics
sales_summary <- data |>
  summarise(
    Mean = mean(sales_revenue),
    Median = median(sales_revenue),
    SD = sd(sales_revenue),
    Min = min(sales_revenue),
    Max = max(sales_revenue),
    Q1 = quantile(sales_revenue, 0.25),
    Q3 = quantile(sales_revenue, 0.75)
  )

sales_summary |>
  kable(caption = "Sales Revenue Summary Statistics ($K)", digits = 2) |>
  kable_styling(bootstrap_options = c("striped", "hover"))
```

```{r sales-distribution, fig.width=10, fig.height=4}
# Sales revenue distribution
p1 <- ggplot(data, aes(x = sales_revenue)) +
  geom_histogram(bins = 30, fill = "#D01012", alpha = 0.7, color = "white") +
  labs(title = "Sales Revenue Distribution", x = "Sales Revenue ($K)", y = "Count") +
  theme_minimal()

p2 <- ggplot(data, aes(y = sales_revenue)) +
  geom_boxplot(fill = "#D01012", alpha = 0.7) +
  labs(title = "Sales Revenue Boxplot", y = "Sales Revenue ($K)") +
  theme_minimal() +
  theme(axis.text.x = element_blank())

gridExtra::grid.arrange(p1, p2, ncol = 2)
```

The average weekly sales revenue is **$`{r} round(mean(data$sales_revenue), 2)`K** with a standard deviation of **$`{r} round(sd(data$sales_revenue), 2)`K**.

## Marketing Spend by Channel

```{r spend-columns}
# Identify marketing spend columns
spend_cols <- names(data)[str_detect(names(data), "_spend$") &
                          !str_detect(names(data), "total_marketing_spend")]
spend_cols
```

```{r spend-summary}
# Summary statistics for all marketing channels
spend_summary <- data |>
  select(all_of(spend_cols)) |>
  pivot_longer(everything(), names_to = "Channel", values_to = "Spend") |>
  mutate(Channel = str_remove(Channel, "_spend") |> str_replace_all("_", " ") |> str_to_title()) |>
  group_by(Channel) |>
  summarise(
    Mean = mean(Spend),
    Median = median(Spend),
    SD = sd(Spend),
    Min = min(Spend),
    Max = max(Spend),
    Total = sum(Spend)
  ) |>
  arrange(desc(Mean))

spend_summary |>
  kable(caption = "Marketing Spend Summary by Channel ($K)", digits = 2) |>
  kable_styling(bootstrap_options = c("striped", "hover"))
```

```{r spend-distribution, fig.width=12, fig.height=8}
# Distribution of spend by channel
data |>
  select(all_of(spend_cols)) |>
  pivot_longer(everything(), names_to = "Channel", values_to = "Spend") |>
  mutate(Channel = str_remove(Channel, "_spend") |> str_replace_all("_", " ") |> str_to_title()) |>
  ggplot(aes(x = Spend, fill = Channel)) +
  geom_histogram(bins = 20, alpha = 0.8, color = "white") +
  facet_wrap(~Channel, scales = "free", ncol = 2) +
  labs(title = "Marketing Spend Distribution by Channel",
       x = "Weekly Spend ($K)", y = "Count") +
  theme(legend.position = "none") +
  scale_fill_brewer(palette = "Set2")
```

## Holiday Week Patterns

```{r holiday-summary}
# Holiday week frequency
holiday_summary <- data |>
  count(holiday_week) |>
  mutate(
    Percentage = n / sum(n) * 100,
    holiday_week = factor(holiday_week, levels = c(0, 1), labels = c("Regular Week", "Holiday Week"))
  )

holiday_summary |>
  kable(caption = "Holiday Week Distribution", digits = 1,
        col.names = c("Week Type", "Count", "Percentage (%)")) |>
  kable_styling(bootstrap_options = c("striped", "hover"))
```

```{r holiday-sales, fig.width=10, fig.height=5}
# Sales comparison: holiday vs non-holiday weeks
data |>
  mutate(holiday_week = factor(holiday_week, levels = c(0, 1),
                               labels = c("Regular Week", "Holiday Week"))) |>
  ggplot(aes(x = holiday_week, y = sales_revenue, fill = holiday_week)) +
  geom_boxplot(alpha = 0.7) +
  labs(title = "Sales Revenue: Holiday vs Regular Weeks",
       x = "Week Type", y = "Sales Revenue ($K)") +
  theme(legend.position = "none") +
  scale_fill_manual(values = c("#0055BF", "#D01012"))
```

```{r holiday-stats}
# Statistical comparison
holiday_stats <- data |>
  mutate(holiday_week = factor(holiday_week, levels = c(0, 1),
                               labels = c("Regular Week", "Holiday Week"))) |>
  group_by(holiday_week) |>
  summarise(
    Mean_Sales = mean(sales_revenue),
    Median_Sales = median(sales_revenue),
    SD = sd(sales_revenue),
    N = n()
  )

holiday_stats |>
  kable(caption = "Sales Statistics by Week Type ($K)", digits = 2) |>
  kable_styling(bootstrap_options = c("striped", "hover"))
```

**Holiday weeks show `{r} ifelse(holiday_stats$Mean_Sales[2] > holiday_stats$Mean_Sales[1], "higher", "lower")` average sales ($`{r} round(holiday_stats$Mean_Sales[2], 2)`K) compared to regular weeks ($`{r} round(holiday_stats$Mean_Sales[1], 2)`K).**

# Temporal Patterns

## Sales Revenue Trends

```{r sales-trends, fig.width=12, fig.height=6}
# Sales over time
ggplot(data, aes(x = date, y = sales_revenue)) +
  geom_line(color = "#D01012", linewidth = 1) +
  geom_smooth(method = "loess", se = TRUE, color = "#0055BF", linewidth = 1.5) +
  labs(title = "Sales Revenue Over Time",
       x = "Date", y = "Sales Revenue ($K)") +
  theme_minimal()
```

## Marketing Spend Trends by Channel

```{r spend-trends, fig.width=12, fig.height=8}
# Spend trends by channel
data |>
  select(date, all_of(spend_cols)) |>
  pivot_longer(-date, names_to = "Channel", values_to = "Spend") |>
  mutate(Channel = str_remove(Channel, "_spend") |> str_replace_all("_", " ") |> str_to_title()) |>
  ggplot(aes(x = date, y = Spend, color = Channel)) +
  geom_line(linewidth = 0.8) +
  facet_wrap(~Channel, scales = "free_y", ncol = 2) +
  labs(title = "Marketing Spend Trends by Channel",
       x = "Date", y = "Weekly Spend ($K)") +
  theme(legend.position = "none") +
  scale_color_brewer(palette = "Set2")
```

## Seasonal Patterns

```{r quarterly-patterns, fig.width=12, fig.height=5}
# Quarterly patterns
data |>
  group_by(quarter) |>
  summarise(
    Avg_Sales = mean(sales_revenue),
    Avg_Total_Spend = mean(total_marketing_spend),
    .groups = "drop"
  ) |>
  pivot_longer(-quarter, names_to = "Metric", values_to = "Value") |>
  mutate(Metric = recode(Metric,
                        "Avg_Sales" = "Average Sales",
                        "Avg_Total_Spend" = "Average Marketing Spend")) |>
  ggplot(aes(x = quarter, y = Value, fill = Metric)) +
  geom_col(position = "dodge", alpha = 0.8) +
  labs(title = "Quarterly Sales and Marketing Spend Patterns",
       x = "Quarter", y = "Amount ($K)", fill = "") +
  scale_fill_manual(values = c("#D01012", "#0055BF")) +
  theme(legend.position = "top")
```

```{r monthly-patterns, fig.width=12, fig.height=5}
# Monthly patterns
data |>
  group_by(month) |>
  summarise(
    Avg_Sales = mean(sales_revenue),
    N = n(),
    .groups = "drop"
  ) |>
  ggplot(aes(x = factor(month), y = Avg_Sales)) +
  geom_col(fill = "#D01012", alpha = 0.8) +
  labs(title = "Average Sales Revenue by Month",
       x = "Month", y = "Average Sales Revenue ($K)") +
  theme_minimal()
```

# Marketing Channel Analysis

## Total Spend by Channel

```{r channel-totals, fig.width=10, fig.height=6}
# Total spend by channel
channel_totals <- data |>
  select(all_of(spend_cols)) |>
  summarise(across(everything(), sum)) |>
  pivot_longer(everything(), names_to = "Channel", values_to = "Total_Spend") |>
  mutate(
    Channel = str_remove(Channel, "_spend") |> str_replace_all("_", " ") |> str_to_title(),
    Percentage = Total_Spend / sum(Total_Spend) * 100
  ) |>
  arrange(desc(Total_Spend))

ggplot(channel_totals, aes(x = reorder(Channel, Total_Spend), y = Total_Spend, fill = Channel)) +
  geom_col(alpha = 0.8) +
  geom_text(aes(label = paste0(round(Percentage, 1), "%")),
            hjust = -0.2, size = 4) +
  coord_flip() +
  labs(title = "Total Marketing Spend by Channel",
       x = "", y = "Total Spend ($K)") +
  theme(legend.position = "none") +
  scale_fill_brewer(palette = "Set2") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))
```

```{r channel-table}
channel_totals |>
  kable(caption = "Total Marketing Spend by Channel", digits = 2,
        col.names = c("Channel", "Total Spend ($K)", "Percentage (%)")) |>
  kable_styling(bootstrap_options = c("striped", "hover"))
```

## Channel Spend Correlations

```{r spend-correlation, fig.width=10, fig.height=8}
# Correlation matrix of marketing channels
spend_cor <- data |>
  select(all_of(spend_cols)) |>
  rename_with(~str_remove(., "_spend") |> str_replace_all("_", " ") |> str_to_title()) |>
  cor()

corrplot(spend_cor, method = "color", type = "upper",
         addCoef.col = "black", number.cex = 0.8,
         tl.col = "black", tl.srt = 45,
         col = colorRampPalette(c("#D01012", "white", "#0055BF"))(200),
         title = "Marketing Channel Spend Correlations",
         mar = c(0, 0, 2, 0))
```

```{r spend-cor-table}
# Show correlation matrix as table
spend_cor |>
  as.data.frame() |>
  rownames_to_column("Channel") |>
  kable(caption = "Marketing Channel Spend Correlation Matrix", digits = 3) |>
  kable_styling(bootstrap_options = c("striped", "hover"))
```

# Relationship Analysis

## Sales vs. Total Marketing Spend

```{r sales-vs-total, fig.width=10, fig.height=6}
# Sales vs total marketing spend
ggplot(data, aes(x = total_marketing_spend, y = sales_revenue)) +
  geom_point(alpha = 0.6, color = "#0055BF") +
  geom_smooth(method = "lm", se = TRUE, color = "#D01012", linewidth = 1.5) +
  labs(title = "Sales Revenue vs. Total Marketing Spend",
       x = "Total Marketing Spend ($K)", y = "Sales Revenue ($K)") +
  theme_minimal()
```

```{r overall-correlation}
# Calculate correlation
total_cor <- cor(data$total_marketing_spend, data$sales_revenue)
cat("Correlation between total marketing spend and sales revenue:", round(total_cor, 3))
```

## Sales vs. Individual Marketing Channels

```{r channel-correlations, fig.width=12, fig.height=10}
# Correlations between individual channels and sales
channel_cors <- data |>
  select(all_of(spend_cols), sales_revenue) |>
  summarise(across(all_of(spend_cols), ~cor(., sales_revenue))) |>
  pivot_longer(everything(), names_to = "Channel", values_to = "Correlation") |>
  mutate(Channel = str_remove(Channel, "_spend") |> str_replace_all("_", " ") |> str_to_title()) |>
  arrange(desc(Correlation))

# Display correlation table
channel_cors |>
  kable(caption = "Correlation between Marketing Channels and Sales Revenue", digits = 3) |>
  kable_styling(bootstrap_options = c("striped", "hover"))

# Visualize correlations
ggplot(channel_cors, aes(x = reorder(Channel, Correlation), y = Correlation, fill = Correlation)) +
  geom_col(alpha = 0.8) +
  geom_text(aes(label = round(Correlation, 3)), hjust = -0.2, size = 4) +
  coord_flip() +
  labs(title = "Correlation: Marketing Channels vs. Sales Revenue",
       x = "", y = "Correlation Coefficient") +
  scale_fill_gradient2(low = "#D01012", mid = "white", high = "#0055BF",
                       midpoint = 0, limits = c(-1, 1)) +
  theme(legend.position = "none") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)), limits = c(0, 1))
```

```{r channel-scatterplots, fig.width=12, fig.height=10}
# Scatterplots of each channel vs sales
data |>
  select(all_of(spend_cols), sales_revenue) |>
  pivot_longer(-sales_revenue, names_to = "Channel", values_to = "Spend") |>
  mutate(Channel = str_remove(Channel, "_spend") |> str_replace_all("_", " ") |> str_to_title()) |>
  ggplot(aes(x = Spend, y = sales_revenue)) +
  geom_point(alpha = 0.5, color = "#0055BF") +
  geom_smooth(method = "lm", se = TRUE, color = "#D01012", linewidth = 1) +
  facet_wrap(~Channel, scales = "free_x", ncol = 2) +
  labs(title = "Sales Revenue vs. Individual Marketing Channels",
       x = "Weekly Spend ($K)", y = "Sales Revenue ($K)") +
  theme_minimal()
```

# Data Quality

```{r data-quality}
# Check for missing values
total_missing <- sum(is.na(data))

missing_by_column <- data |>
  summarise(across(everything(), ~sum(is.na(.)))) |>
  pivot_longer(everything(), names_to = "Column", values_to = "Missing_Count") |>
  mutate(Percentage = Missing_Count / nrow(data) * 100) |>
  filter(Missing_Count > 0) |>
  arrange(desc(Missing_Count))

if (nrow(missing_by_column) > 0) {
  missing_by_column |>
    kable(caption = "Missing Values by Column", digits = 2) |>
    kable_styling(bootstrap_options = c("striped", "hover"))
} else {
  cat("No missing values detected in the dataset.\n")
}

cat("\nTotal missing values across all columns:", total_missing, "\n")
```

# Key Findings

## Summary of Insights

1. **Sales Performance**
   - Average weekly sales: $`r round(mean(data$sales_revenue), 2)`K
   - Sales range from $`r round(min(data$sales_revenue), 2)`K to $`r round(max(data$sales_revenue), 2)`K

2. **Marketing Investment**
   - Total marketing spend: $`r round(sum(data$total_marketing_spend), 0)`K over the analysis period
   - Top spending channel: `r channel_totals$Channel[1]` (`r round(channel_totals$Percentage[1], 1)`% of budget)
   - Channel spend shows varying levels of correlation with each other

3. **Holiday Effect**
   - Holiday weeks represent `r round((sum(data$holiday_week) / nrow(data)) * 100, 1)`% of observations
   - Sales during holiday weeks are `r ifelse(holiday_stats$Mean_Sales[2] > holiday_stats$Mean_Sales[1], paste0(round((holiday_stats$Mean_Sales[2] - holiday_stats$Mean_Sales[1]) / holiday_stats$Mean_Sales[1] * 100, 1), "% higher"), paste0(round((holiday_stats$Mean_Sales[1] - holiday_stats$Mean_Sales[2]) / holiday_stats$Mean_Sales[2] * 100, 1), "% lower"))` than regular weeks

4. **Sales-Marketing Relationships**
   - Overall correlation between total marketing spend and sales: `r round(total_cor, 3)`
   - Strongest performing channel (by correlation): `r channel_cors$Channel[1]` (r = `r round(channel_cors$Correlation[1], 3)`)
   - All channels show `r ifelse(all(channel_cors$Correlation > 0), "positive", "mixed")` relationships with sales

5. **Data Quality**
   - Dataset is complete with `r ifelse(total_missing == 0, "no missing values", paste0(total_missing, " missing values"))`
   - Temporal coverage: `r n_distinct(data$week)` weeks across `r n_distinct(data$year)` year(s)

## Implications for Modeling

Based on this exploratory analysis, the following considerations are important for Marketing Mix Modeling:

1. **Feature Engineering**
   - **Adstock transformations** should be applied to capture carryover effects of marketing spend
   - **Saturation curves** are needed to model diminishing returns at higher spend levels
   - **Holiday indicator** should be included as a control variable

2. **Model Considerations**
   - The positive correlations between channels and sales support causal modeling approaches
   - Time-series aspects suggest using temporal train/test splits rather than random splits
   - The presence of seasonality indicates the need for time-based controls

3. **Channel Prioritization**
   - Channels with stronger correlations may warrant more sophisticated transformation parameters
   - Budget optimization should consider both historical performance and diminishing returns

4. **Next Steps**
   - Implement adstock and saturation transformations (as done in `model.R`)
   - Compare multiple modeling approaches (frequentist vs. Bayesian)
   - Validate models using appropriate time-series cross-validation
   - Conduct sensitivity analysis on transformation parameters

---

## Session Information

```{r session-info}
sessionInfo()
```
