---
title: "Marketing Mix Model Analysis"
subtitle: "Model Diagnostics and Next Steps"
author: "MMM Analysis"
date: today
format:
  html:
    code-fold: true
    code-summary: "Show code"
    toc: true
    toc-depth: 3
    toc-location: left
    theme: cosmo
    embed-resources: true
execute:
  warning: false
  message: false
---

# Executive Summary

This document provides a comprehensive analysis of four Marketing Mix Models (MMM):

1. **Simple Linear Regression**: Baseline model with raw features
2. **Transformed Linear Regression**: Adstock + saturation transformations
3. **Bayesian LR (Transformed)**: Bayesian with transformations
4. **Bayesian LR (Raw)**: Bayesian without transformations

```{r setup}
#| include: false
library(tidyverse)
library(tidymodels)
library(rstanarm)
library(patchwork)
library(scales)
library(bayesplot)

# Load data
data <- read_csv("data/synthetic-marketing-sales.csv", show_col_types = FALSE)

# Load saved models
model1 <- readRDS("saved_models/model1_simple_lr.rds")
model2 <- readRDS("saved_models/model2_transformed_lr.rds")
model3 <- readRDS("saved_models/model3_bayes_transformed.rds")
model4 <- readRDS("saved_models/model4_bayes_raw.rds")
comparison <- readRDS("saved_models/comparison_results.rds")
params <- readRDS("saved_models/transformation_params.rds")
metadata <- readRDS("saved_models/metadata.rds")

train_data <- comparison$train_data_split
test_data <- comparison$test_data_split
```

# Actual vs. Predicted Analysis

```{r actual-vs-predicted-setup}
# Get predictions for all models on test data
model1_test_pred <- predict(model1$fit, test_data) %>%
  bind_cols(test_data %>% select(date, sales_revenue)) %>%
  mutate(model = "1. Simple LR")

model2_test_pred <- predict(model2$fit, model2$test_data) %>%
  bind_cols(model2$test_data %>% select(date, sales_revenue)) %>%
  mutate(model = "2. Transformed LR")

model3_test_pred <- tibble(
  date = test_data$date,
  sales_revenue = model3$test_data %>%
    mutate(sales_revenue = test_data$sales_revenue) %>%
    pull(sales_revenue),
  .pred = predict(model3$fit, newdata = model3$test_data)
) %>%
  mutate(model = "3. Bayesian (Transformed)")

model4_test_pred <- tibble(
  date = test_data$date,
  sales_revenue = model4$test_data %>%
    mutate(sales_revenue = test_data$sales_revenue) %>%
    pull(sales_revenue),
  .pred = predict(model4$fit, newdata = model4$test_data)
) %>%
  mutate(model = "4. Bayesian (Raw)")

all_predictions <- bind_rows(
  model1_test_pred,
  model2_test_pred,
  model3_test_pred,
  model4_test_pred
)
```

```{r actual-vs-predicted-plot, fig.width=12, fig.height=8}
ggplot(all_predictions, aes(x = date)) +
  geom_line(aes(y = sales_revenue, color = "Actual"), linewidth = 1) +
  geom_line(aes(y = .pred, color = "Predicted"), linewidth = 0.8, alpha = 0.7) +
  facet_wrap(~model, ncol = 2, scales = "free_y") +
  scale_color_manual(
    values = c("Actual" = "#2E86AB", "Predicted" = "#A23B72"),
    name = ""
  ) +
  scale_y_continuous(labels = label_comma()) +
  labs(
    title = "Actual vs. Predicted Sales Revenue",
    subtitle = "Test set performance comparison across all models",
    x = "Date",
    y = "Sales Revenue ($)",
    caption = paste0("Test period: ", min(test_data$date), " to ", max(test_data$date))
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "top",
    plot.title = element_text(face = "bold", size = 16),
    strip.text = element_text(face = "bold"),
    panel.grid.minor = element_blank()
  )
```

## Scatter Plots: Predicted vs. Actual

```{r scatter-plots, fig.width=12, fig.height=8}
ggplot(all_predictions, aes(x = sales_revenue, y = .pred)) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "gray40") +
  geom_point(alpha = 0.6, size = 2.5, color = "#2E86AB") +
  geom_smooth(method = "lm", se = TRUE, color = "#A23B72", linewidth = 0.8) +
  facet_wrap(~model, ncol = 2, scales = "free") +
  scale_x_continuous(labels = label_comma()) +
  scale_y_continuous(labels = label_comma()) +
  labs(
    title = "Predicted vs. Actual Sales (Test Set)",
    subtitle = "Perfect predictions would lie on the diagonal line",
    x = "Actual Sales Revenue ($)",
    y = "Predicted Sales Revenue ($)"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    strip.text = element_text(face = "bold"),
    panel.grid.minor = element_blank()
  )
```

## Recommended Next Steps

1. **Short Term**
   - Implement channel-specific adstock parameters
   - Add interaction terms between channels
   - Test positive-constrained priors

2. **Medium Term**
   - Collect additional variables (competitor data, economic indicators)
   - Implement hierarchical model for seasonal patterns
   - Add lagged dependent variable (with appropriate caution)

3. **Long Term**
   - Set up automated model retraining pipeline
   - Implement A/B testing framework for marketing spend
   - Build budget optimization tool based on model insights
